---
title: "`ggplot2` Shortcuts"
author: "Jonathan Carroll"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{`ggplot2` Shortcuts}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This package allows easy access to some common `ggplot2` tasks. For example, rotating the `x` axis labels is a very frequently looked up task, and we can make it easier. If we create a simple `ggplot2` plot

```{r, message = FALSE}
library(ggplot2)
library(cowplot)
library(ggeasy)

p <- ggplot(mtcars, aes(hp, mpg)) + geom_point()
```

then by default, this looks like

```{r, fig.width = 6, fig.height = 6}
p + labs(title = "ggplot2 default")
```

We can perform various rotations though

```{r, fig.width = 8, fig.height = 8}
p1 <- p + 
    easy_rotate_x_labels() + 
    labs(title = "default rotation")
p2 <- p + 
    easy_rotate_x_labels(angle = 45, side = "right") + 
    labs(title = "angle = 45")
p3 <- p + 
    easy_rotate_x_labels("startattop") + 
    labs(title = "text starts at top")
p4 <- p + 
    easy_rotate_x_labels("startatbottom") + 
    labs(title = "text starts at bottom")
    
plot_grid(p1, p2, p3, p4, nrow = 2, align = "hv", axis = "l")
```

## Removing legends

Removing legends is made easier by the `easy_remove_legend` function.  When called without arguments, all legends are removed (equivalent to `theme(legend.position = "none")`). Alternatively, the names of aesthetics for which legends should be removed can be passed.

```{r, fig.width = 8, fig.height = 8}
p <- ggplot(mtcars, aes(wt, mpg, colour = cyl, size = hp)) +
    geom_point()

p1 <- p + 
    labs(title = "With all legends")
p2 <- p + 
    easy_remove_legend() + 
    labs(title = "Remove all legends")
p3 <- p + 
    easy_remove_legend(size) + 
    labs(title = "Remove size legend")
p4 <- p + 
    easy_remove_legend(size, color) + 
    labs(title = "Remove both legends specifically")

plot_grid(p1, p2, p3, p4, nrow = 2, align = "hv", axis = "l")
```

## The `ggbot`

Sometimes you don't even remember these helper functions. `ggbot` takes your commands and tries to interpret them in a way that modifies a plot. 

```{r, fig.width = 8, fig.height = 8}
p <- ggplot(mtcars, aes(cyl, hp, col = "red")) + geom_point()

p1 <- p + 
    labs(title = "Default red points")
p2 <- p + 
    ggbot("blue points") + 
    labs(title = "A string request")
p3 <- p + 
    ggbot("points", "blue") +
    labs(title = "The order is irrelevant")
p4 <- p + 
    ggbot("blue", "point") +
    labs(title = "Partial match to words")
    
plot_grid(p1, p2, p3, p4, nrow = 2, align = "hv", axis = "l")
```

Various combinations of aesthetics can be requested, for example we can ask to change all colours of points

```{r, fig.width = 5, fig.height = 5}
ggplot(mtcars, aes(wt, mpg, colour = factor(cyl), size = hp)) + 
    geom_point() + 
    theme_dark() + 
    ggbot("make points red white and blue")
```

`ggbot` can also be used to remove components easily

```{r, fig.width = 8, fig.height = 8}
p <- ggplot(mtcars, aes(cyl, hp, col = "red")) + geom_point()

p1 <- p + 
    labs(title = "Default red points")
p2 <- p + 
    ggbot("drop x axis") + 
    labs(title = "No x axis")
p3 <- p + 
    ggbot("y axis remove") +
    labs(title = "No y axis")
p4 <- p + 
    ggbot("remove legend") +
    labs(title = "No legend")
    
plot_grid(p1, p2, p3, p4, nrow = 2, align = "hv", axis = "l")
```

or remove everything `theme`-related

```{r, fig.width = 5, fig.height = 5}
ggplot(mtcars, aes(wt, mpg, colour = factor(cyl), size = hp)) + 
    geom_point() + 
    ggbot("I don't need no theme") +
    ggbot("I don't need no legend")
```


